---
title: "pd_distribution"
author: "Yaqian Huang"
date: "2023-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(lmerTest)


```

## Load, tidy, and code the dataset

The dataset includes the counts of pd and fry - two types of voicing. These counts are logged from different phrasal positions in sentences with a fixed carrier and different sentence-medial compound stimuli.

The DV is the count data; the IVs are different lexical tones, phrasal positions, and types of voicing. Random variables include subject, and the interaction between subject and each of the IVs.

The following sections include the analyses of the distribution of the counts as a function of tone in the compound stimuli and as a function of phrasing and tone across the entire sentence, and the plots showing the distributions.

The list of at most five analysis or modeling issues that are of primary concern:

-   specification of random-effect structure supported by data (convergence issues if include all; added optimizer)

-   model selection between a nested random effects and varying slopes and intercepts

-   plots of interactions of partial effects between tone and phrasing on count

-   model diagnostics

-   power statistics

```{r loaddf, warning=F,message=F}

pros <- read_csv("data1.csv")

# filter out derived neutral tone in sentence-medial stimuli
pros <- pros[!(pros$Tone=='0'&pros$Phrase%in%c('PM','PI','PF')),]

pros <- pros %>% mutate(id = str_sub(Filename,start=5,end=6),
               gender = ifelse(str_sub(Filename,start=3,end=3)=="0",'Women','Men'))

pros$ind <- paste0(ifelse(pros$gender=='Women','f','m'),pros$id)

pros$group[pros$Phrase%in%c('PM','PI','PF')] <- 'Medial'
pros$group[pros$Phrase%in%c('UI','PS1','PS2')] <- 'Initial'
pros$group[pros$Phrase%in%c('AS2','AS1','UF','AS')] <- 'Final'

pros$Phrase[pros$Phrase%in%c('AS2','AS1','AS')] <- 'AS'

pros$Phrase <- factor(pros$Phrase,
                      levels = c('UI','PS2','PS1','PI','PM','PF','AS','UF'))

pros$Type[pros$Type=='pd'] = 'period doubling'
pros$Type[pros$Type=='fry'] <- 'vocal fry'

pros$group <- factor(pros$group, levels = c("Initial", "Medial", "Final"))

pros_style <- list(
  scale_color_manual(values = c('darkorange1','cornflowerblue','azure4')),#BBBBBB
  scale_fill_manual(values = c('darkorange1','cornflowerblue','azure4')),#009988
  theme_light(base_size = 20),
  theme(
    aspect.ratio     = .75,
    legend.text = element_text(size = 22),
    strip.background = element_blank(),
    strip.text       = element_text(color = "black",size=18, hjust = 0),
    strip.placement = 'inside',
    panel.border     = element_rect(color = "black", fill = NA),
    legend.position = "bottom",
    legend.title = element_text(size=22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size=18)
  )
)
```

## Tone models sentence-medially

```{r tmodels, cache=T, warning=F,message=F}

tone.by.ph <- pros %>% filter(Phrase%in%c('PI','PM','PF'))%>%
  group_by(Type,Phrase,Tone,ind)%>%summarise(count=n())

tone.by.ph$Tone <- factor(tone.by.ph$Tone, levels = c("1", "2", "3",'4'), 
                    labels = c("T1", "T2", "T3",'T4'))

tone.by.ph$Type <- factor(tone.by.ph$Type, levels = c( "period doubling", "vocal fry"))

tm.pd <- glmer(count~Type*(Tone+Phrase)+Tone:Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Type)+(1|ind:Phrase)
               ,data=tone.by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))
drop1(tm.pd,test='Chisq',trace = T)
summary(tm.pd)

#control=lmerControl(optimizer="nloptwrap", optCtrl=list(xtol_abs=1e-8, ftol_abs=1e-8)))

tm.pd2 <- glmer(count~Type*Tone*Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Type)+(1|ind:Phrase)
               ,data=tone.by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))
summary(tm.pd2)
drop1(tm.pd2,test='Chisq',trace = T)
anova(tm.pd,tm.pd2)

# tm.pd3 <- glmer(count~Type*Tone*Phrase+(1|ind)+(Tone|ind)+(Type|ind)+(Phrase|ind)
#                ,data=tone.by.ph,family = 'poisson',
#                control=lmerControl(optimizer="bobyqa"))
# summary(tm.pd3)
# tm.pd4 <- glmer(count~Type*(Tone+Phrase)+Tone:Phrase+(1|ind)+(Tone|ind)+(Type|ind)+(Phrase|ind)
#                ,data=tone.by.ph,family = 'poisson',
#                control=lmerControl(optimizer="bobyqa"))
# summary(tm.pd4)

#tm.pd2@optinfo$conv$lme4
#with(tm.pd@optinfo$derivs,max(abs(solve(Hessian,gradient)))<2e-3)
#tm.pd@optinfo$derivs %>% with(. , solve(Hessian, gradient))  %>% abs() %>% max()

tm <- glmer(count~Tone+Phrase+Type+(1|ind)+(1|ind:Tone)+(1|ind:Type)+(1|ind:Phrase),
             data=tone.by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))
drop1(tm,test='Chisq', trace=T)

summary(tm.pd)


# refit the model for vocal fry as default

tone.by.ph$Type <- factor(tone.by.ph$Type, levels = c( "vocal fry", "period doubling"))

tm.vf <- glmer(count~Type*(Tone+Phrase)+Tone:Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Type)+(1|ind:Phrase)
               ,data=tone.by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))

summary(tm.vf)

tm.vf2 <- glmer(count~Type*Tone*Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Type)+(1|ind:Phrase)
               ,data=tone.by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))
summary(tm.vf2)
drop1(tm.vf2,test='Chisq',trace = T)
anova(tm.vf,tm.vf2)



```

## Plot of tonal distributions

```{r tplot, warning=F,message=F}
pros.tone.ph <- pros %>%filter(group=='Medial')%>%group_by(Type)%>%
  mutate(type.n = n())%>%
  filter(group=='Medial')%>%group_by(Type,Tone,Phrase)%>%
  mutate(ph.n = n())%>%group_by(Type,Tone,Phrase)%>%
  summarize(Proportion=ph.n/type.n)%>%unique()

pros.tone.ph$Tone <- factor(pros.tone.ph$Tone, levels = c("1", "2", "3",'4'), 
                    labels = c("Tone 1 (55)", "Tone 2 (35)", "Tone 3 (21[4])",'Tone 4 (51)'))

pros.tone <- pros %>% filter(group=='Medial')
pros.tone$Tone <- factor(pros.tone$Tone, levels = c("1", "2", "3",'4'), 
                    labels = c("Tone 1 (55)", "Tone 2 (35)", "Tone 3 (21[4])",'Tone 4 (51)'))


pros.tone%>%group_by(Type,Tone)%>%
  summarize(n = n())%>%
  mutate(Proportion=prop.table(n))%>%ggplot()+
  geom_histogram(aes(x=Type,y=Proportion,fill=Type,group=Type),size = 1,
           stat = 'identity', position = position_dodge(),
                 alpha=.3,color='black')+ #alpha=0
  facet_grid(~Tone)+pros_style+theme(axis.text.x = element_blank(),aspect.ratio = 1)+ylim(c(0, 0.6))


pros.tone.ph %>% ggplot()+
  geom_bar(aes(x=Phrase,y=Proportion,fill=Type,group=Type),size = 1,
           stat = 'identity', position = position_dodge(),
                 alpha=.3,color='black')+ #alpha=0
  facet_grid(~Tone)+pros_style+theme(aspect.ratio = 1)

```

## Phrasing models across the utterance

```{r pmodels, cache=T, warning=F,message=F}

by.ph <- pros%>%group_by(Type,Phrase,Tone,ind)%>%summarise(count=n())

by.ph$Tone <- factor(by.ph$Tone, levels = c("1", "2", "3",'4','0'), 
                    labels = c("T1", "T2", "T3",'T4','T0'))

by.ph$Type <- factor(by.ph$Type, levels = c( "period doubling", "vocal fry"))

pm.pd <- glmer(count~Tone+Type*Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Phrase)+(1|ind:Type),data=by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))

with(pm.pd@optinfo$derivs,max(abs(solve(Hessian,gradient)))<2e-3)

drop1(pm.pd,test='Chisq',trace = T)

pm <- glmer(count~Tone+Type+Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Phrase)+(1|ind:Type),data=by.ph,family = 'poisson')
drop1(pm,test='Chisq',trace = T)

summary(pm.pd)

# refit the model for vocal fry as default
by.ph$Type <- factor(by.ph$Type, levels = c( "vocal fry", "period doubling"))

pm.vf <- glmer(count~Tone+Type*Phrase+(1|ind)+(1|ind:Tone)+(1|ind:Phrase)+(1|ind:Type),data=by.ph,family = 'poisson',
               control=lmerControl(optimizer="bobyqa"))

summary(pm.vf)


```

## Plot of phrasal distribution

```{r pplots, warning=F,message=F}

pros %>%filter()%>%group_by(Type,Phrase)%>%
  summarize(n = n())%>%
  mutate(Proportion=prop.table(n)) %>% ggplot()+
  geom_bar(aes(x=Phrase,y=Proportion,group=Type,fill=Type),stat = 'identity',alpha=0.3,
           color='black',
                 size=1,
                 position = position_dodge())+
  pros_style+theme(aspect.ratio = .6)+ylim(c(0, 0.5))

pros %>%filter(Type=='period doubling')%>%group_by(Type,Phrase)%>%
    summarize(n = n())%>%
    mutate(Proportion=prop.table(n)) %>% ggplot()+
    geom_bar(aes(x=Phrase,y=Proportion,group=Type,fill=Type),stat = 'identity',alpha=.3,
             color='black',
             size=.8, width=.5,
             position = position_dodge())+pros_style+theme(aspect.ratio = .3)

pros %>%#group_by(Type,Phrase)%>%
  #summarize(n = n())%>%
  #mutate(Proportion=prop.table(n)) %>%
  ggplot()+
  geom_density(aes(x=Phrase,group=Type,fill=Type),stat = 'count',alpha=0.3,
           color='black')+
  pros_style+theme(aspect.ratio = .6)

```

```{r}
pros$Phrase <- as.character(pros$Phrase)

pros$Phrase[pros$Phrase=='PM'] = 'syll2'
pros$Phrase[pros$Phrase=='PI'] = 'syll1'
pros$Phrase[pros$Phrase=='PF'] = 'syll3'
pros$Phrase[pros$Phrase=='UI'] = 'I'
pros$Phrase[pros$Phrase=='PS2'] = 'teach'
pros$Phrase[pros$Phrase=='PS1'] = 'you'
pros$Phrase[pros$Phrase=='AS'] = 'how-to'
pros$Phrase[pros$Phrase=='UF'] = 'say'

pros$Phrase <- factor(pros$Phrase,
                      levels = c('I','teach','you','syll1','syll2','syll3','how-to','say'))


pros %>%filter(Type=='period doubling')%>%group_by(Type,Phrase)%>%
  summarize(n = n())%>%
  mutate(Proportion=prop.table(n)) %>% ggplot()+
  geom_bar(aes(x=Phrase,y=Proportion,group=Type,fill=Type),stat = 'identity',alpha=0.3,
           color='black',
                 size=1, width=.45, just=1,
                 position = position_dodge())+
  pros_style+theme(aspect.ratio = .6)+ylim(c(0, 0.5))

pros %>% group_by(Type,Phrase)%>%
  summarize(n = n())%>%
  mutate(Proportion=prop.table(n)) %>% filter(Type=='vocal fry') %>% ggplot()+
  geom_bar(aes(x=Phrase,y=Proportion,group=Type,fill=Type),stat = 'identity',alpha=0.3,
           color='black',
           size=1,
           width=0.45,just=0,
           position = position_dodge())+  scale_fill_discrete(drop=FALSE) +
  scale_x_discrete(drop=FALSE)+
  pros_style+theme(aspect.ratio = .6)+ylim(c(0, 0.5))
```
